A multi-scale story of the diffusion of a new technology: the web

### 2024-06-18 paper/submission1/tranos2023.tex
postpresentation

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
  pdfauthor={Emmanouil Tranos},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Digital economy in the UK: a multi-scalar story of the diffusion
of web technologies}
\author[1]{Emmanouil Tranos%
\corref{cor1}%
}
 \ead{e.tranos@bristol.ac.uk}

@@

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{sec1}{%
\section{Introduction}\label{sec1}}

Geographers were always interested in how new technologies and
innovations diffuse across space and time and, importantly, how such
spatio-temporal processes can be modelled. After all, diffusion together
with invention and innovation are considered the pillars of
technological change \citep{das2022diffusion}. The seminal contribution
of \citet{hagerstrand1968innovation} is illustrative of this early
interest. However, the torch of exploring and modelling such processes
had been passed to other disciplines such as economics, business studies
and sociology well before the `cultural turn' of economic geography
\citep{perkins2005international}. A potential explanation of the lack of
geographical studies exploring the diffusion of new and, more
specifically for this paper, digital technologies across \emph{both}
space and time can be attributed to the scarcity of relevant and
granular enough data. As \citet{zook2022mapping} highlight, digital
activities are hardly ever captured in official data.

@@

@@

This paper focuses on the diffusion of the web as new technology during
the \(1996-2012\) period. This was an exciting period for digital
technologies as it corresponds with the commercialisation of the
internet and, consequently, its almost universal adoption. The reader is
reminded that it was only in 1994 when Netscape Navigator was
introduced, a year before Microsoft's Internet Explorer \footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}.
Also, only \(9\) per cent of UK's households had access to the internet
in \(1998\) \citep{ons2018}, the web included mostly static webpages,
there were no social media and web browsing was happening exclusively
from desktop PCs as there were no smartphones \citep{tranosuk}. Hence,
it is fair to say that the study period captures the very early stages
of the diffusion on a new technology that is the web. This is key point
in the lifecycle of a new technology. Firstly, during this period new
technologies are expensive, crude and imperfect
\citep{rosenberg1994exploring, wilson201281}. A simple comparison
between Web 1.0 and Web 3.0 applications clearly illustrates this
\citep{tranos2020social}. During this period the performance of a new
technology is the main attraction and not the cost to access and use it
\citep{wilson2011lessons} There is a broader theoretical discussion in
the literature regarding the motivation behind early adoption. As
summarised by \citet{perkins2005international}, on the one hand,
epidemic models highlight the role of interpersonal contacts as a way
for new technologies to diffuse. On the other hand, economic models
underline the importance of heterogeneity. Different firms have
different structures and business plans, which define the potential
economic returns of the adoption of a new technology and, therefore, the
choice to adopt a new technology becomes an individual option. From a
broader and evolutionary perspective, initial conditions are essential
for the creation and evolution of path-dependent technological
development trajectories \citep{neffke2011regions, simmie2014new}. This
argument is even more relevant when the focus is on Information and
Communication Technologies (ICT) because of the commonly found lag
between investment and economic returns as reflected in the Solow
paradox \citep{acemoglu2014return, brynjolfsson2018artificial}.

Importantly, the data used here depicts the \emph{active} engagement
with the web in the UK as it contains all the commercial websites
archived by the Internet Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
that include a mention to at least one valid UK postcode in the web
text. Websites are identified as commercial ones if they are part of the
.co.uk second level domain (SLD). Also, the act of of creating a website
is understood here as active engagement with the web vis-à-vis the more
passive act of browsing the web or having an internet connection
\citep{tranosuk}. Previous studies have focused mostly on more passive
notions of engaging with digital technologies such as internet adoption
and internet speeds
\citep[e.g.][]{blank2018local, riddlesden2014broadband}. More details
about the data and the data generation process can be found in Section
\protect\hyperlink{sec3}{XX}. \textbf{ADD DATA DESCRIPTION, WHY .CO.UK}

@@

\hypertarget{sec2}{%
\section{Literature review}\label{sec2}}

@@

The temporal dimension was further explored by Everett Rogers and his
seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
Rogers being a sociologist, his work focused not on the diffusion of
innovations over space and time, but instead on the adoption of new
technologies and innovations by individuals and the individual
mechanisms that drive the decisions behind adoption. He identified five
groups of individuals regarding their adoption speed: innovators, early
adopters, early majority, late majority and laggards. Key mechanism of
diffusion and adoption is always communication and how knowledge is
transferred within a social system. Therefore, all approaches agree on
the S-shaped diffusion and cumulative adoption pattern
\citep{grubler1990rise}.

Schmidt's Law empirically illustrates a similar pattern. \emph{Core} and
usually highly agglomerated regions is where new technologies are
invented and commercially deployed {[}\citet{grubler1990rise};
grubler1990rise{]}. This is where the first adopters tend to be based.
Then, technologies spread to the \emph{rim} and eventually to the
\emph{periphery}. Although adoption pace might be higher when new
technologies finally arrive to the periphery, the saturation levels
there may never reach the ones in the core because of the lack of
infrastructure or other necessary institutions
\citep{leibowicz2016representing}. \citet{grubler1990rise} effectively
summarises the three key characteristics of the spatial diffusion
process: (i) the cumulative level of adoption follows an S-shaped
pattern just like purely temporal models; (ii) diffusion is shaped by a
hierarchy effect in a form of a centrifugal force: from core to
periphery; and (iii) diffusion is also shaped by distance and a
neighbourhood effect and contaminate nearby locations.

The remaining of this section reviews empirical studies which analysed
the diffusion on new technologies over space and time. Although the
spatial dimension is present in most of the following studies, the level
of spatial detail is almost always more coarse than the one adopted
here. \citet{beardsell1999spatial} studied the evolution of the computer
industry in \(317\) US metro areas during the \(1977-1992\) period using
employment data. Their analysis indicated that the relative size
distribution holds for urban computer employment and also urban
heterogeneity is essential in explaining this distribution. in a recent
study, \citet{bednarz2020pulled} focused on wind turbines and modelled
their spatial diffusion across \(402\) German regions during
\(1970-2015\). Their key finding is that local demand than local supply
was the main driving factor.

At a global scale \citet{perkins2005international} explored whther the
diffusion rate of new technologies is driven by a latecomer advantage
and the engagement with the global economy via foreign direct
investments and trade. Their results illustrate that indeed latecomers
and developing countries experience diffusion of new technologies more
rapidly than early adopter and developed countries. At the same scale,
\citet{perkins2011internet} explored whether the adoption of previous
communication technologies that is mail, telegrams and telephones was
shaped by similar socioeconomic factors as the internet. Their results
indicated common patterns regarding the drivers behind the adoption of
different communication technologies.

@@

\hypertarget{sec3}{%
\section{Materials and Methods}\label{sec3}}

For instance, we can approximate digital economic activities in a region
by measuring the volume of commercial webpages by considering only
.co.uk websites, which are dedicated to commercial activities (Thelwall,
2000). Such commercial websites are used to exchange information,
support online transactions and share opinions (Blazquez \& Domenech,
2018). Although nothing prevents a UK-based company from adopting a
generic TLD such as.com and, indeed, such cases escape our data, we do
not expect that such omissions could affect our results given the
popularity of the .uk ccTLD: UK consumers prefer to visit a .uk website
when they are searching for products or services (Hope, 2017); and
anecdotal evidence indicates that during the first half of 2000, three
.co.uk domains were registered every minute (OECD, 2001). Also, as Table
1 demonstrates, .co.uk was the most popular SLD under the .uk ccTLD in
2000. Hence, we use the total volume of all the archived webpages in a
region as a proxy for the level of digitization. We start the analysis
by focusing on commercial webpages (.co.uk).

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{old-ideas}{%
\subsection{Old ideas}\label{old-ideas}}

Hexagon density maps: reflect the spatial structure of Britain. Websites
are associated in places where people live and work.

\textbf{TODO} maps at the local authority level per firms and discuss
patterns

Neighbourhood effect: diffusion proceeds outwards from innovation
centers, first ``hitting'' nearby rather than far-away locations
(Grubler 1990)

\begin{itemize}
\item
  Moran's I: for OA and LAD over time
\item
  LISA maps: for OA and LAD over time More and less expected clusters.
  Different scales show different results
\item
  Website density regressions: for OA and LAD over time \textbf{TODO}
  add 0s Similar pattern
\end{itemize}

Hierarchy effect: from main centers to secondary ones -- central places

S-shaped diffusion curves

\begin{itemize}
\item
  S for LADs per firm and OA.
\item
  Fast and slow LAs map. There is clustering
\item
  s\_uk\_firm\_11
\item
  ranks: there is stability and movement
\end{itemize}

\hypertarget{new-structure}{%
\subsection{New structure}\label{new-structure}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples
\item
  ranks: there is stability and movement
\item
  Neighbourhood effect: diffusion proceeds outwards from innovation
  centers, first ``hitting'' nearby rather than far-away locations
  (Grubler
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1989}
\tightlist
\item
\end{enumerate}

\begin{itemize}
\item
  Moran's I: for OA and LAD over time
\item
  LISA maps: for OA and LAD over time More and less expected clusters.
  Different scales show different results
\item
  Website density regressions: for OA and LAD over time
\end{itemize}

@@

\hypertarget{rf-results}{%
\subsubsection{RF results}\label{rf-results}}

The next section incorporates the above discussed spatial drivers of the
diffusion of web technologies into the same modelling framework. The aim
is to use variables depicting these spatial processes in order to
predict the diffusion of web technologies in the UK over space and time
and across different scales. Specifically, four different models are
estimated. Firstly, all the data points at the OA and LAD are utilised
in order to build RF models and assess their capacity to predict the
adoption of web technologies. These two models will reveal the
predictive capacity of the spatial processes and also allow to see how
the importance of such variables changes across scales. The next two
models will be again trained on web difusion at the two working scales:
OA and LAD. However, instead of using all the data points, the OA and
the LAD from one of the 12 UK regions are hold out and then the trained
model is used to predict web adoption in the OA or the LAD of the
held-out region. This process takes place recursively for all UK
regions. The difference in the predictive capacity of the different
samples will reveal how different are these spatial process across
regions and, importantly, at different scales.

It needs to be highlighted here that the cross-validation for all models
presented here is spatially and temporally sensitive. Instead of using
10 random folds, we employ the \texttt{CAST} package which allows holds
back data points from specific years and spatial units and use them for
testing in order to estimate the model performance
\citep{meyer2018improving}.

The models need to include variables that capture the three processes
that the relevant literature and the descriptive analysis presented in
the previous section highlighted. Namely, the models capture: (i) a
hierarchy effect with diffusion running from main centres to secondary
ones, (ii) a neighborhood effect according to which diffusion first hits
nearby locations, and (iii) the rather canonical pattern of the
diffusion over time as reflected in the S-shaped pattern in the
cumulative level of adoption.

To capture the hierarchy effect the models include as predictors a one
year lag of website density in London -- the largest city in the UK, a
one year lag of the website density in the nearest city and the same for
the nearest retail centre. Due to the small sizes of the retail centres,
the latter is only relevant for the OA-level models. In addition, the
models include the distance to London, the nearest city and the nearest
retail centre. The underlying logic is that the level of website
adoption in a spatial unit depends on the level of the adoption in
places further up in the urban hierarchy the previous year. To depict
the neighbourhood effect, the web density of the neighbouring spatial
units in the previous year is calculated. Again, the underpinning
rationale is that the level of web adoption within a spatial unit
depends on the level of web adoption in the neighbouring spatial units
the year before. This is the `hitting nearby locations first' argument.
Therefore, the spatial and temporal lag of the website density in LAD
and OA is calculated. Lastly, the time effect which is reflected on the
S curve for the cumulative adoption is captured by trend variable.
Hence, all four model will follow the following generic form
\ref{model}:

@@

Table \ref{table.metrics.all} presents the model performance for the
first set on models, for which all data points are employed for training
and testing via cross validation. The first one is trained and tested on
374 LAD and the second on 232,296 OA for a 16 year period (1997-2012).
The results are remarkably good considering that the are the outcome of
space and time sensitive CV, so the the model does not suffer from
overfit. At the LAD level we the model predicts 81\% of the variation of
website density. Both error metrics indicate that the model error is
\((\frac{1}{20} - \frac{1}{30})\) of a website per firm. At the OA this
score drops down to 21\%. Considering its granularity, this is still a
remarkable performance. To contextualise it, the model results in a MAE
of one website for areas small enough to host less than 140
households.\footnote{According to the Office for National Statistics,
  80\% of OA in England and Wales host 110-139 households,
  \href{https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas}{www.ons.gov.uk}.}
Because of the small size of the spatial units, the distribution is
highly skewed and a significant part of them is not linked to any
websites. In 1997 only 1\% of the UK OA were associated with a website.
This should not come as a surprise as this was the very beginning of the
commercial internet and any activities with a digital footprint were
concentrated in a handful of areas. This was clearly illustrated in
Section \textbf{ADD}. At the end of the study period almost half of the
UK OA were not associated with a website. Again, given the granularity
of the data this should not come as a surprise. Despite the large
heterogeneity in website adoption, the model reached an R-Squared of
0.2.

@@

\end{figure}

\begin{figure}

@@

\end{figure}

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A multi-scale story of the diffusion of a new technology: the web},
  pdfauthor={Emmanouil Tranos},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{A multi-scale story of the diffusion of a new technology: the
web}
\author[1]{Emmanouil Tranos%
\corref{cor1}%
}
 \ead{e.tranos@bristol.ac.uk}

@@

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{sec1}

Geographers were always interested in how new technologies and
innovations diffuse across space and time and, importantly, how such
spatio-temporal processes can be modeled. After all, diffusion together
with invention and innovation are considered the pillars of
technological change \citep{das2022diffusion}. The seminal contribution
of \citet{hagerstrand1968innovation} is illustrative of this early
interest. However, the torch of exploring and modelling such processes
had been passed to other disciplines such as economics, business studies
and sociology well before the `cultural turn' of economic geography
\citep{perkins2005international}. A potential explanation of the lack of
geographical studies exploring the diffusion of new and, more
specifically for this paper, digital technologies across both space and
time can be attributed to the scarcity of relevant and granular enough
data. As \citet{zook2022mapping} highlight, digital activities are
hardly ever captured in official data.

@@

\textbf{SAY SOMETHING ABOUT SPATIAL ANALYTICS, DIST PAPER BY ARP, NOT
SPECIFIC EXPLANATORY VARIABLES, INTANGIBLE}

@@

This paper focuses on the diffusion of the web as new technology during
the \(1996-2012\) period. This was an exciting period for digital
technologies as it corresponds with the commercialisation of the
internet and, consequently, its almost universal adoption. The reader is
reminded that it was only in 1994 when Netscape Navigator was
introduced, a year before Microsoft's Internet Explorer.\footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}
Also, only \(9\) per cent of UK's households had access to the internet
in \(1998\) \citep{ons2018}, the web included mostly static webpages,
there were no social media and web browsing was happening exclusively
from desktop PCs as there were no smartphones \citep{tranosuk}. Hence,
it is fair to say that the study period captures the very early stages
of the diffusion on a new technology that is the web as well as its
maturity. The former is a key point in the lifecycle of a new
technology. Firstly, during this period new technologies are expensive,
crude and imperfect \citep{rosenberg1994exploring, wilson201281}. A
simple comparison between Web 1.0 and Web 3.0 applications clearly
illustrates this \citep{tranos2020social}; for instance a static website
compared with a platform like \texttt{github}, which enables cooperation
between users and the creation of new information, meaning, and
knowledge \citep{faraj2016special, barassi2012does}. During this period
the performance of a new technology is the main attraction and not the
cost to access and use it \citep{wilson2011lessons}. There is a broader
theoretical discussion in the literature regarding the motivation behind
early adoption. As summarised by \citet{perkins2005international}, on
the one hand, epidemic models highlight the role of interpersonal
contacts as a way for new technologies to diffuse. On the other hand,
economic models underline the importance of heterogeneity. Different
firms have different structures and business plans, which define the
potential economic returns of the adoption of a new technology and,
therefore, the choice to adopt a new technology becomes an individual
option. From a broader and evolutionary perspective, initial conditions
are essential for the creation and evolution of path-dependent
technological development trajectories
\citep{neffke2011regions, simmie2014new}. This argument is even more
relevant when the focus is on digital technologies because of the
commonly found lag between investment and economic returns as reflected
in the Solow paradox
\citep{acemoglu2014return, brynjolfsson2018artificial}.

Importantly, the data used here depicts the \emph{active} engagement
with the web in the UK as it contains all commercial websites that (i)
are part of the UK's relevant second level domain (SLD, .co.uk), (ii)
have been archived by the Internet Archive\footnote{\href{See\%20https://archive.org/}{https://archive.org/}.},
and (iii) include a mention to at least one valid UK postcode in the web
text. The act of creating a website is understood here as active
engagement with the web vis-à-vis the more passive act of browsing the
web or having an internet connection \citep{tranosuk}. Previous studies
have focused mostly on more passive notions of engaging with digital
technologies such as internet adoption and internet speeds
\citep[e.g.][]{blank2018local, riddlesden2014broadband}. More details
about the data and the data generation process can be found in Section
\hyperref[sec3]{XX}.

@@

\section{Literature review}\label{sec2}

@@

The temporal dimension was further explored by Everett Rogers and his
seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
Rogers being a sociologist, his work focused not on the diffusion of
innovations over space and time, but instead on the adoption of new
technologies and innovations by individuals and the individual
mechanisms that drive the decisions behind adoption. He identified five
groups of individuals regarding their adoption speed: innovators, early
adopters, early majority, late majority and laggards. The key mechanism
of diffusion and adoption is communication and how knowledge is
transferred within a social system. Therefore, all approaches agree on
the S-shaped diffusion and cumulative adoption pattern
\citep{grubler1990rise}.

Schmidt's Law empirically illustrates a similar pattern. \emph{Core} and
usually highly agglomerated regions is where new technologies are
invented and commercially deployed \citep{grubler1990rise}. This is
where the first adopters tend to be based. Then, technologies spread to
the \emph{rim} and eventually to the \emph{periphery}. Although adoption
pace might be higher when new technologies finally arrive to the
periphery, the saturation levels there may never reach the ones in the
core because of the lack of infrastructure or other necessary
institutions \citep{leibowicz2016representing}. \citet{grubler1990rise}
effectively summarises the three key characteristics of the spatial
diffusion process: (i) the cumulative level of adoption follows an
S-shaped pattern just like purely temporal models; (ii) diffusion is
shaped by a hierarchy effect in a form of a centrifugal force: from core
to periphery; and (iii) diffusion is also shaped by distance and a
neighbourhood effect and contaminate nearby locations. These are the
three mechanisms that the empirical analysis in Section
\phantomsection\label{sec4}{4} investigates.

The remaining of this section reviews empirical studies which analysed
the diffusion on new technologies over space and time. Although the
spatial dimension is present in most of the following studies, the level
of spatial detail is always more coarse than the one adopted here.
\citet{beardsell1999spatial} studied the evolution of the computer
industry in \(317\) US metro areas during the \(1977-1992\) period using
employment data. Their analysis indicated that the relative size
distribution holds for urban computer employment and also urban
heterogeneity is essential in explaining this distribution. In a recent
study, \citet{bednarz2020pulled} focused on wind turbines and modelled
their spatial diffusion across \(402\) German regions during
\(1970-2015\). Their key finding is that local demand than local supply
was the main driving factor.

At a global scale \citet{perkins2005international} explored whther the
diffusion rate of new technologies is driven by a latecomer advantage
and the engagement with the global economy via foreign direct
investments and trade. Their results illustrate that indeed latecomers
and developing countries experience diffusion of new technologies more
rapidly than early adopters and developed countries. At the same scale,
\citet{perkins2011internet} explored whether the adoption of previous
communication technologies that is mail, telegrams and telephones was
shaped by similar socioeconomic factors as the internet. Their results
indicated common patterns regarding the drivers behind the adoption of
different communication technologies.

@@

All in all, \ldots{}

\section{Materials and Methods}\label{sec3}

\textbf{ADD FAMILY OF WEB TECHNOLOGIES}

To capture the diffusion of web technologies, a website density metric
is developed for two different geographical scales: the Local Authority
Districts (LAD) and the Output Areas (OA). The former is an
administrative unit and there are
\textbackslash textcolor\{red\}c.~380\textbackslash textcolor\{black\}
such units in the UK. The latter is a census-based geographical unit,
which is very small as there are c.~230,000 of them in the UK. This
methodological choice will allow the mapping of the diffusion of web
technologies and the assessment of the diffusion mechanisms at these two
very different spatial scales.

The population of websites is calculated using data from the Internet
Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
The Internet Archive is one of the most complete and oldest archive of
webpages in the world operating since 1996
\citep{ainsworth2011much, holzmann2016dawn}. It is a web crawler, which
discovers webpages by following the hyperlinks of every webpage its
archives. This dataset, which is curated by the British Library,
contains all the archived webpages from the UK ccTLD (.uk) from the
1996--2012 period. In essence, this is a long list of 2.5 billion URLs
of archived webpages including also the archival timestamp.

Instead of using the whole .uk country code top-level domain (ccTLD),
this paper focuses on its commercial subset, the .co.uk SLD. This choice
decrease the heterogeneity of the web data as such commercial websites
have rather specific aims: they are used to diffuse information, support
online transactions and share opinions
\citep{THELWALL2000441, blazquez2018big}. Although a UK company can
adopt a generic TLD such as.com and such cases escape the data used
here, such omissions should not affect our results given the popularity
of the .uk ccTLD \citep{tranosuk}: UK consumers prefer to visit a .uk
website when they are searching for products or services \citep{hope};
and anecdotal evidence indicates that during the first half of 2000,
three .co.uk domains were registered every minute \citep{oecd_coms}.
Importantly, previous studies illustrated that .co.uk is the most
popular UK SLD \citep{tranosuk}.

The text from these webpages was scanned using regular expressions
(regex) to identify strings of text which resemble UK postcodes and one
fifth of them included a mention to a postcode \citep{BL2013geo}. This
information allows the geolocation of the data and the creation of the
LAD and OA counts.

Before that, the data cleaning process included a data aggregation step,
through which the webpages were aggregated to the relevant websites.
Based on the following example, all three webpages listed below are part
of the same website (\url{http://www.website.co.uk}) and at the end only
websites and not the nested webpages were considered as otherwise the
metric would have been biased towards large websites.

\begin{itemize}
\item
  \url{http://www.website.co.uk/webpage_a} B152TT
\item
  \url{http://www.website.co.uk/webpage_b} BS81TH
\item
  \url{http://www.website.co.uk/webpage_c} B152TT
\end{itemize}

\noindent What is challenging is that this aggregation approach, which
has been used elsewhere \citep{tranosuk, shoreditch} may lead to
websites with multiple postcodes. As per the above example,
(\url{http://www.website.co.uk}) includes two unique postcodes: B152TT
and BS81SS. The distribution of postcodes for 2000 is presented in Table
\ref{f2000}, which clearly illustrates the wide range. On one hand, we
have websites anchored to a unique location (72\% of all the
reconstructed websites), which may represent a small company with a
single trading location. On the other hand, we have websites with
thousands of different postcodes. Considering the timeperiod of the
analysis, such cases can represent directories which used to be popular
in the pre-search engines early times of internet
\citep{tranosuk}.\footnote{See Figure A1 in \textbf{Appendix A in the
  supplemental data online for examples}.}

The analysis presented here is based on two subsets of these data.
Firstly, the analysis is based on websites which only contain only one
unique postcode and, therefore, the geolocation process might suffer
from less noise. As a robustness check, the analysis is replicated the
analysis for an extended subset of websites, which include up to ten
unique postcodes. These websites are geolocated by equally attaching
them to multiple locations. This extended sample includes 94 percent of
all the archived websites in 2000.

\begin{table}

\caption{\label{tab:unnamed-chunk-1}Number of unique postcodes per .co.uk website, 2000.\label{f2000}}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Postcodes & F & F (\%) & Cummulative F\\
\midrule
(0,1] & 41,596 & 0.718 & 0.718\\
(1,2] & 6,451 & 0.111 & 0.830\\
(2,10] & 6,163 & 0.106 & 0.936\\
(10,100] & 2,975 & 0.051 & 0.988\\
(100,1000] & 646 & 0.011 & 0.999\\
\addlinespace
(1000,10000] & 62 & 0.001 & 1.000\\
(10000,100000] & 4 & 0.000 & 1.000\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Source: } Tranos et al. 2021}\\
\end{tabular}
\end{table}

To create the web density metric, the yearly website counts at the LAD
level are standardised by the number of firms in LAD to avoid biases
associated with LADs hosting a large number of firms. Given that there
is no such statistic for the OA, the actual OA level counts are used.
However, because of the the rather consistent spatial definition of OA
(they host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
website counts in OA are interpreted as a density metric too.

This website metric\ldots{}

@@

\section{Results}\label{sec4}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples
\item
  ranks: there is stability and movement
\item
  Neighbourhood effect: diffusion proceeds outwards from innovation
  centers, first ``hitting'' nearby rather than far-away locations
  (Grubler 1990)
\end{enumerate}

\begin{itemize}
\item
  Moran's I: for OA and LAD over time
\item
  LISA maps: for OA and LAD over time More and less expected clusters.
  Different scales show different results
\end{itemize}

@@

\subsubsection{RF results}\label{rf-results}

The next section incorporates the above discussed spatial processes of
the diffusion of web technologies into the same modelling framework. The
aim is to use variables depicting these spatial processes in order to
predict the diffusion of web technologies in the UK over space and time
and across different scales. Specifically, four different models are
estimated. Firstly, all the data points at the OA and LAD are utilised
in order to build RF models and assess their capacity to predict the
adoption of web technologies. These two models will reveal the
predictive capacity of the spatial processes and also allow to see how
the importance of such variables changes across scales. The next two
sets of models will be again trained on web diffusion at the two working
scales: OA and LAD. However, instead of using all the data points, the
OA and the LAD from one of the twelve UK regions are held out and then
the trained model is used to predict web adoption in the OA or the LAD
of the held-out region. This process takes place recursively for all UK
regions. The difference in the predictive capacity of the different
samples will reveal how dissimilar are these spatial process across
regions and, importantly, at different scales.

It needs to be highlighted here that the cross-validation for all models
is spatially and temporally sensitive. Instead of using 10 random folds,
we employ the \texttt{CAST} package which allows holding back data
points from specific years and spatial units and use them for testing in
order to estimate the model performance \citep{meyer2018improving}.

The models need to include variables that capture the three processes
that the relevant literature and the descriptive analysis presented in
the Section \textbf{XX} highlighted. Namely, the models capture: (i) a
hierarchy effect with diffusion running from main centres to secondary
ones, (ii) a neighborhood effect according to which diffusion first hits
nearby locations, and (iii) the rather canonical pattern of diffusion
over time as reflected in the S-shaped pattern in the cumulative level
of adoption.

To capture the hierarchy effect the models include as predictors a one
year lag of website density in London -- the largest city in the UK, a
one year lag of the website density in the nearest city and the same for
the nearest retail centre. Due to the small sizes of the retail centres,
the latter is only relevant for the OA-level models. In addition, the
models include the distance to London, the nearest city and the nearest
retail centre. The underlying logic is that the level of website
adoption in a spatial unit depends on the level of the adoption in
places further up in the urban hierarchy the previous year. To depict
the neighbourhood effect, the web density of the neighbouring spatial
units in the previous year is employed. Again, the underpinning
rationale is that the level of web adoption within a spatial unit
depends on the level of web adoption in the neighbouring spatial units
the year before. This is the `hitting nearby locations first' argument.
Therefore, the spatial and temporal lag of the website density in LAD
and OA is calculated. Lastly, the time effect which is reflected on the
S curve for the cumulative adoption is captured by trend variable.
Hence, all four model will follow the following generic form (Eq.
\ref{model}):

@@

Table \ref{table.metrics.all} presents the model performance for the
first set on models, for which all data points are employed for training
and testing via cross validation. The first one is trained and tested on
374 LAD and the second on 232,296 OA, both for a 16 year period
(1997-2012). The results are remarkably good considering that the are
the outcome of space and time sensitive CV, so the the model does not
suffer from overfitting. At the LAD level the model predicts 81\% of the
variation of website density. Both error metrics indicate that the model
error is \((\frac{1}{20}, \frac{1}{30})\) of a website per firm. At the
OA the RSquared drops down to 21\%. Considering its granularity, this is
still a remarkable performance. To contextualise it, the model results
in a MAE of one website for areas small enough to host less than 140
households.\footnote{According to the Office for National Statistics,
  80\% of OA in England and Wales host 110-139 households,
  \href{https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas}{www.ons.gov.uk}.}
Because of the small size of the spatial units, the distribution is
highly skewed and a significant part of them is not linked to any
websites. In 1997 only 1\% of the UK OA were associated with a website.
This should not come as a surprise as this was the very beginning of the
commercial internet and any activities with a digital footprint were
concentrated in a handful of areas. This was clearly illustrated in
Section \textbf{ADD}. At the end of the study period almost half of the
UK OA were not associated with a website. Again, given the granularity
of the data this should not come as a surprise.

@@

\end{figure}%

\begin{figure}[H]

@@

\end{figure}%
